### 2024-04-09 Chapters/Chapter4.tex
Uppfært eftir ábendingar ÓPP
Aðeins uppfært eftir ath frá KJ 9/4/24

\section{Model structure}
The structure of the neural network is such that it contains some number $n$ of fully connected layers and batch normalization for each layer, along with regularization. All of these layers have the same number of units. The last layer has a dropout of 50\%. In addition to these layers there is one more output layer. This is simply a dense layer with 1 unit. A grid search was performed to determine the hyperparameters that minmize the loss. Hyperparameters are parameters that are set before the training begins\cite{hyperparameter_definition}. These hyperparameters include number of units in each layer, number of epochs to train for, number of layers, batch size, optimizer and penalty to enforce in the regularization. The possible combinations tried can be seen in Table \ref{table:gridSearchHyperparamters}.

@@

\section{Feature selection and importance}
Using the three datasources, a model is trained. The data from Veðurstofa only represents the ground truth and is not used as part of the training data. The gust ($f_g$) and wind speed ($f$) are used to calculate the target gust factor. All weather parameters used in predictions come from the Carra reanalysis data, with variables wind speed, wind direction, temperature, pressure. All of these variables are queried in three height levels (15, 250, 500 meters above ground) at the location of a given weather station. The second data source for the training data, is the elevation data. Veðurstofa provided a GeoTif file that contains elevation information above sea level, covering Iceland, in a 20 by 20 grid. A couple of different landscape distributions were looked at. A sector upwind, a sector upwind and downwind as well as a circle surrounding a weather station.

@@

Using all the variables from Carra, along with the two derived factors ($Ri$, $N$) and two sectors (upwind and downwind) downscaled to 10 features using PCA, the feature importance can be seen in Figure \ref{fig:ShapleyWaterfallFirstTry}. The waterfall graph in Figure \ref{fig:ShapleyWaterfallFirstTry} shows that the Richardson number for higher levels is most important.

>>>>

\section{Model structure}
The structure of the neural network is such that it contains some number $n$ of fully connected layers and batch normalization for each layer, along with regularization. All of these layers have the same number of units. The last layer has a dropout of 50\%. In addition to these layers there is one more output layer. This is simply a dense layer with 1 unit. A grid search was performed to determine the hyperparameters that minmize the loss. Hyperparameters are parameters that are set before the training begins\cite{hyperparameters_definition}. These hyperparameters include number of units in each layer, number of epochs to train for, number of layers, batch size, optimizer and penalty to enforce in the regularization. The possible combinations tried can be seen in Table \ref{table:gridSearchHyperparamters}.

@@

\section{Feature selection and importance}
Using the three datasources, a model is trained. The data from IMO only represents the ground truth and is not used as part of the training data. The gust ($f_g$) and wind speed ($f$) are used to calculate the target gust factor. All weather parameters used in predictions come from the CARRA reanalysis data, with variables wind speed, wind direction, temperature, pressure. All of these variables are queried in three height levels (15, 250, 500 meters above ground) at the location of a given weather station. The second data source for the training data, is the elevation data. IMO provided a GeoTif file that contains elevation information above sea level, covering Iceland, in a 20 by 20 grid. A couple of different landscape distributions were looked at. A sector upwind, a sector upwind and downwind as well as a circle surrounding a weather station.

@@

Using all the variables from CARRA, along with the two derived factors ($Ri$, $N$) and two sectors (upwind and downwind) downscaled to 10 features using PCA, the feature importance can be seen in Figure \ref{fig:ShapleyWaterfallFirstTry}. The waterfall graph in Figure \ref{fig:ShapleyWaterfallFirstTry} shows that the Richardson number for higher levels is most important.

### 2024-04-09 MSc.tex
Aðeins uppfært eftir ath frá KJ 9/4/24

% Packages
\usepackage[utf8]{inputenc}
\usepackage[icelandic, english]{babel}
\usepackage{t1enc}
\usepackage{graphicx}
\usepackage[intoc]{nomencl}
\usepackage{enumerate,color}
\usepackage{url}
\usepackage[pdfborder={0 0 0}]{hyperref}
\BeforeTOCHead[toc]{\cleardoublepage\pdfbookmark{\contentsname}{toc}} % Add Table of Contents to PDF "bookmark" table of contents
\usepackage{appendix}
\usepackage{eso-pic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[sf,normalsize]{subfigure}
\usepackage[format=plain,labelformat=simple,labelsep=colon]{caption}
\usepackage{placeins}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{xspace}

@@

@@

>>>>

% Packages
\usepackage[utf8]{inputenc}
\usepackage[icelandic, english]{babel}
\usepackage{t1enc}
\usepackage{graphicx}
\usepackage[intoc]{nomencl}
\usepackage{enumerate,color}
\usepackage{url}
\usepackage[pdfborder={0 0 0}]{hyperref}
\BeforeTOCHead[toc]{\cleardoublepage\pdfbookmark{\contentsname}{toc}} % Add Table of Contents to PDF "bookmark" table of contents
\usepackage{appendix}
\usepackage{eso-pic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[sf,normalsize]{subfigure}
\usepackage[format=plain,labelformat=simple,labelsep=colon]{caption}
\usepackage{placeins}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{xspace}
\usepackage{float}

@@

% https://tex.stackexchange.com/questions/374661/keyword-highlighting-not-working-in-python-with-listings
\lstdefinestyle{Python}{
    language        = Python,
    basicstyle      = \ttfamily,
    keywordstyle    = \color{blue},
    keywordstyle    = [2] \color{teal}, % just to check that it works
    stringstyle     = \color{green},
    commentstyle    = \color{red}\ttfamily
}

@@

% This must be placed after "\begin{document}":
\lstset{
    frame       = single,
    numbers     = left,
    showspaces  = false,
    showstringspaces    = false,
    captionpos  = t,
    caption     = \lstname
}
